---
title: "HR_Analytics"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
## Read data in R  

```{r cars}
hr_data <- read.csv("/Users/prateek/Desktop/01.2 Coursework/01_Statistical Computing/Proj/WA_Fn-UseC_-HR-Employee-Attrition.csv")
head(data)
```

## Data Structure   

```{r}
str(hr_data)
```
### Data Summary  

```{r}
summary(hr_data)
```

### checking duplictes   
### Checking NA   

```{r}
anyDuplicated(hr_data)  #0 duplicates
colSums(is.na(hr_data))
```

## Mean (Avg) Age by Gender  

```{r}
library(dplyr)
avg_age <- hr_data %>% select(Gender, Age) %>% group_by(Gender) %>% summarize(avg=mean(Age))
avg_age

```


## Attrition split in Dataset  

```{r}
library(ggplot2)
distrinbution <-  table(hr_data$Attrition)
addmargins(distrinbution)
round(prop.table(distrinbution), digits = 2)
```

## Attriton based on Gender  

```{r}
plot1a <- ggplot(hr_data) +
        geom_bar(mapping = aes(x = Gender,fill = Attrition))
plot1a

```
### No significant realtionship in Attrition based on Gender  

### Attrition seems to be equally likely based on gender  


## Total working years VS Attrition  


```{r}
plot1 <- ggplot(hr_data) +
        geom_bar(mapping = aes(x = TotalWorkingYears,fill = Attrition))

plot1 + scale_fill_manual(breaks = c("No", "Yes"), 
                       values=c("springgreen4", "red2"))

```

## Distance from Home VS Attrition  
```{r}
plot2 <- ggplot(hr_data) +
        geom_bar(mapping = aes(x = DistanceFromHome,fill = Attrition),position = "dodge")

plot2 + scale_fill_manual(breaks = c("No", "Yes"), 
                       values=c("springgreen4", "red2"))
```

## Attrition -> Total Working Years and Age  
```{r}
plot2_1 <-  ggplot(hr_data,aes(x=TotalWorkingYears,y=Age,color=Attrition)) +               
            geom_boxplot()+ggtitle("Attrition -> Total Working Years and Age")

plot2_1
```

## Attrition based on Monthly income and Age  

```{r}
plot3 <-  ggplot(hr_data,aes(x=Age,y = MonthlyIncome, color=Attrition)) + 
          geom_point()+ggtitle("Attrition -> Monthly income and Age")

plot3
```

```{r}
plot4 <-  ggplot(hr_data,aes(x=Age,y=MonthlyIncome,fill=Attrition)) +               geom_boxplot()+ggtitle("Attrition -> Monthly income and Age")

plot4 + scale_fill_manual(breaks = c("No", "Yes"), 
                       values=c("springgreen4", "red2"))
```

### As expected Younger generation has low income  

### Also, attrition rate is higher with younger generation  


## Below Plot represents higher attrition rate because of low Job Satisfaction  

```{r}
plot6 <-  ggplot(hr_data,aes(x=Attrition,  y =JobSatisfaction,fill=Attrition)) +        geom_boxplot()+ggtitle("Job Satisfaction vs Attrition")

plot6
```

## We observe low attrition rate in population with Mritial Status as Married  
```{r}
plot7 <-  ggplot(data=hr_data, aes(x=MaritalStatus, fill = Attrition)) +
        geom_bar() + 
        ggtitle("Attrition based on Maritial Status")

plot7
```

## We observe high attrition ratio in Departments-> Human Resources & Technical Degree  
```{r}
plot8 <- ggplot(hr_data, aes(x = EducationField, fill = Attrition)) + 
        geom_bar(position = "dodge") +
        ggtitle("Attrition based on Departments")
plot8 + scale_fill_manual(breaks = c("No", "Yes"), 
                       values=c("springgreen4", "red2"))

```




```{r}
head(hr_data)
```
## Attrition based on Job Satisfaction  
```{r}
plot9 <-  ggplot(data=hr_data, aes(x=JobSatisfaction, fill = Attrition)) +
        geom_bar(position = "dodge")
        ggtitle("Attrition based on Job Satisfaction")
        
plot9 + scale_fill_manual(breaks = c("No", "Yes"), 
                       values=c("springgreen4", "red2"))
```


## Attrition based on Work Life Satisfaction  
```{r}

plot10 <-  ggplot(data=hr_data, aes(x=WorkLifeBalance, fill = Attrition)) +
        geom_bar(position = "dodge")
        ggtitle("Attrition based on Work Life Satisfaction")
        
plot10 + scale_fill_manual(breaks = c("No", "Yes"), 
                       values=c("springgreen4", "red2"))
```

## Attrition based on Loyalty Index


```{r}
hr_data <- hr_data %>%
      mutate(loyalty_index = (TotalWorkingYears/ (NumCompaniesWorked + 1)))
     
plot11 <- ggplot(hr_data,aes(x=loyalty_index,y = Attrition, color=Attrition)) +
          geom_point()+ggtitle("Attrition based on Loyalty Index")
plot11
```

```{r}
plot12 <-  ggplot(hr_data,aes(x=WorkLifeBalance,y = JobSatisfaction, color=Attrition)) + geom_point()+ggtitle("Attrition -> Monthly income and Age")


plot12 + scale_fill_manual(breaks = c("No", "Yes"), 
                       values=c("springgreen4", "red2"))
```













